---
title: "Homework 1"
author: "Emer Gunne"
date: today
date-format: long
format: pdf
fontsize: 1.2em
editor: source
echo: false
---
## Clustering on Iris dataset
The Iris dataset in R programming contains 150 observations of 4 variables:

*Sepal Length* & *Sepal Width* , *Petal Length* & *Petal Width*

The 5th Column contains the **Species** label for each observation of which there are three:
**Setosa** , **Versicolor** and **Virginica**

__View pairs of clusters:__

I would like to view a matrix of the 4 variables showing how we can see clusters based on the dimensions of the variables which we can then relate to species

```{r}
data("iris") #load R package dataset Iris 
Cluster <- iris[,-5] #take out species column,use sepal length & width
#and petal length & width to form pairwise clusters of the data
pairs(Cluster, gap = 0, pch = 19, col = adjustcolor ( "blue", alpha.f = 0.4),
      main = "Clusters")#use the function pairs to form a matrix of the Cluster dataset 
#use a circle for each datapoint and use blue with a lighter tone for contrast. 
```

### Descriptive statistics

I will now investigate the variation in the dataset using summary statistics

**All Species**

```{r}
summary(iris)
```

**Setosa**

```{r}
setosa <- subset (iris, Species =="setosa")# create a subset of the data for species setosa
versicolor <- subset (iris, Species =="versicolor")# create a subset of the data for species versicolor
virginica <- subset (iris, Species =="virginica")#create a subset of the data for species virginiCA
summary(setosa)# look at the variation in the data for setosa species
```

**Versicolor**

```{r}
summary(versicolor)
```

**Virginica**

```{r}
summary(virginica)
```

### In summary

Species Setosa has the shortest sepals and petals, widest sepals and narrowest petals.

Species Versicolor has the narrowest sepals.

Species Virginica has the longest sepals and petals.

Petal length and petal width set the species Setosa apart as can be seen in the cluster matrix.
